### 2018-12-11 Papers/SHADE/elsarticle-template/paperSHADE.tex
SHADE added to paper

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{chngpage}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\modulolinenumbers[5]

\journal{Applied Soft Computing}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Elsevier\fnref{myfootnote}}
\address{Radarweg 29, Amsterdam}
\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
\ead[url]{www.elsevier.com}

\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{support@elsevier.com}

\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

\section{Background}

\section{Constrained Clustering}

Clustering is the task of grouping instances of a dataset into subsets. The criterion used to assign an instance to a given cluster is the similarity with the rest of elements in that cluster, and the dissimilarity with the rest of instances of the dataset. In this way, the clustering process obtains subsets of instances of the initial data set that present high intra-similarity and low inter-similarity.

In most clustering applications it is common to have some kind of information about the data set to be analyzed. In constrained clustering this information is given in the form of pairs of instances. A constraint states whether the instances which it refers to must, or must not, be assigned to the same cluster. Using this type of information it is possible to obtain a better result than by using completely unsupervised clustering algorithms.

Before formalizing the definition of constraints we must give a notation for the data set. We will note our dataset as $X$, which is a set of $N$ instances with $D$ features. We well refer to the instances of $X$ as $x$. Now, given the definition of dataset, we can define two types of constraints:

\begin{itemize}

	\item Must-link constraints $c_=(x_j,x_i)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.

	\item Cannot-link constraints $c_{\neq}(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ cannot be assigned to the same cluster.

\end{itemize}

The goal of constrained clustering is to find a partition (or clustering) of $K$ clusters $C = {c_1 \cdots c_K}$ of the dataset $X$ that ideally satisfies all constraints in the constraints set. As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $N = |X| = \sum_{i = 1}^{K} |c_i| $.

Knowing how a constraint is defined, Must-link constraints---from now on ML---are an example of an equivalence relation; therefore, CL constraints are reflexive, transitive and symmetric. This way, given the constraints $c_=(x_a,x_b)$ and $c_=(x_b,x_c)$ then $c_=(x_a,x_c)$ is verified. In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $c_=(x_a,x_b)$, then $c_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{xu2013improving}\cite{davidson2007survey}.

On the other hand, Cannot-link restrictions---from now on CL---do not constitute an equivalence relation, but they can be entailed. To illustrate that, we consider the instances $x_a \in c_i$ and $x_b \in c_j$, and the constraint $c_{\neq}(x_a,x_b)$, then it is also true that $c_{\neq}(x_c,x_d)$ for all $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.

\textcolor{red}{Comentar la complejidad del problema?}

\section{Background}

\section{Brief Review of BRKGA Algorithm}

The biased random-key genetic algorithm---from now on BRKGA---was first proposed by Gonçalves and Resende \cite{gonccalves2011biased} as a generalization of the random-key genetic algorithm \cite{bean1994genetic}.

In BRKGA each solution---also called chromosome---is represented as a vector of values in the interval $[0,1]$. However, this vector is not a solution to the problem, it is only a representation of it. This allows us to abstract the details of the problem to apply a genetic algorithm that operates with real coding. Therefore, a deterministic decoder is necessary to obtain the actual solution to the problem based on the chromosome that encodes it.

BRKGA also requires a fitness function to evaluate each chromosome. This function takes the decoded solution as input and provides its fitness value as output. It is designed specifically for each problem.

The BRKGA optimization process is completely independent of the problem to solve. First, a population $P$ of $p_{size}$ vectors of random-keys is initialized, this will be the initial generation. Each random-key vector has $N$ random-keys in it. The population is sorted by the fitness value $f_i$ of each chromosome to select the first $p_e$ individuals, this is, the elite of the population. The elite will be preserved in the next generation without modification. To introduce diversity, a number $p_m$ of new random-keys are also included in the next generation. The remaining chromosome of the new generation ($p_{size} - p_e - p_m$) are obtained through crossovers between elite and non-elite parents \cite{de2017comparison}.

\subsection{Adaptation of BRKGA for Constrained Clustering} \label{sec:AdaptationofBRKGA}

Oliviera, Chaves and Lorena (2017) proposed an adaptation of BRKGA to apply it to constrained clustering \cite{de2017comparison}.

Their method randomly initializes a population of random-key vectors, each one with $N$ random-keys. The decoder divides the interval $[0,1]$ in $K$ intervals, so there exist a correspondence between each random-key and the integer corresponding to the interval in which it is. Table \ref{tab:table1} shows an example of random-key decodification for $K = 3$. Note that extreme values 0 and 1 can also appear in a random-key vector.

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{0.9}
	\resizebox{\textwidth}{!}{
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|}
		\hline
		Index & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\ 
		\hline
		Random-key & 0.12 & 0.37 & 0.66 & 0.56 & 0.00 & 0.97 & 0.23 & 0.25 & 0.15 & 1.00 \\
		\hline
		Clusters & 1 & 2 & 2 & 2 & 1 & 3 & 1 & 3 & 1 & 3 \\ 
		\hline

	\end{tabular}}
	\caption{Random-key decodification example \cite{de2017comparison}}
	\label{tab:table1}
\end{table}

Once the decoded solution has been obtained, the fitness value of the solution is computed as in Equation \eqref{eq1}.

\begin{equation}
f_i = z_i + \overbracket{(\mu * N * {infeasibility}_i)}^{penalty}
\label{eq1}
\end{equation}

Where $\mu$ is a high value, $infeasibility_i$ is the number of non-satisfied constraints and $z_i$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq2}.

\begin{equation}
z_i = \frac{\sum_{x_a, x_b \in c_i} d_D^2(x_a,x_b)}{|c_i|} \;\;\; \forall c_i \in C
\label{eq2}
\end{equation}

Where $x_a$ and $x_b$ are instances in the cluster $c_i$ such that $a \neq b$ and the distance between each pair of instances in $c_i$ is included in the sum only once.

Oliviera, Chaves and Lorena also added a new element to BRKGA, which is a local search. This local optimization method is applied to each decoded chromosome in each generation, but its results are not transferred to the chromosome, in order to maintain the diversity. So the results of the local search are only used to update the best solution found so far if needed.

\section{The SHADE Optimization Method}

Success history based adaptive differential evolution---from now on SHADE---is an optimization method based on differential evolution. Proposed by Tanabe and Fukunaga (2013), it was an improvement for the JADE model, created by Zhang and Sanderson \cite{tanabe2013success}\cite{zhang2009jade}.

SHADE uses a historical record of successful parameters as a mechanism for the adaptation of parameters that are involved in the process of creating new generations. Unlike JADE, which, in order to create a new generation, only considers the parameters used in the previous generation.

\subsection{SHADE Operators}

As in differential evolution, SHADE uses muntant generation, crossover and replacement operators to explore the space of solutions and bring in new generations of individuals.

The mutation strategy used by SHADE is known as current-to-pbest/1, and the expression that generates new individuals is described in the Equation \eqref{eq3}.

\begin{equation}
m_{i,G} = p_{i,G} + F_i * (p_{pbest, G} - p_{i,G}) + F_i * (p_{r1, G} - p_{r2,G})
\label{eq3}
\end{equation}

Where $m_{i,G}$ is the mutant vector, which is generated based on an individual from the population $p_{i,G}$. The indices $r1$ and $r2$ are random values in the range $[0,N]$, different from $i$ and from each other. The individual $p_{pbest, G}$ is randomly selected from among the best $N \times p\;\;|\;\;p\in [0,1]$ individuals in the population. In this way the $p$ parameter controls the greediness of the mutation strategy. The parameter $F_i$ defines the magnitude of the operator.

After generating the mutant vector $m_{i,G}$, a crossover operator is applied to combine it with the parent $p_{i,G}$ vector, the result is the trial vector. SHADE uses the Binomial crossover operator, which is defined by the Equation \eqref{eq4}.

\begin{equation}
t_{j,i,G} = \left\{ \begin{array}{lc}
m_{j,i,G} &   if \;\; rand[0,1) \le CR_i \;\; or \;\;j = j_{rand} \\
p_{j,i,G} &  otherwise
\end{array}
\right.
\label{eq4}
\end{equation}

Where $rand[0,1)$ is a random number selected from a normal distribution in the range $[0,1)$, $j_{rand}$ is a random integer selected from the range $[1,D]$, and $CR \in [0,1]$ is the crossover ratio.

The SHADE replacement operator determines the survivors individuals for the next generation. It compares each parent $x_{i,G}$ with the trial vector $_{j,i,G}$, generated based on it. The best one of these two individuals will survive for the next generation. Equation \eqref{eq5} shows this concept.

\begin{equation}
p_{i,G + 1} = \left\{ \begin{array}{lc}
t_{i,G} &   if \;\; f(t_{i,G}) \le f(p_{i,G}) \\
p_{i,G} &  otherwise
\end{array}
\right.
\label{eq5}
\end{equation}

To maintain diversity, SHADE can make use of an external archive $A$, in which parents $p_{i,G}$ who were replaced by their associated trial vector $t_{i,G}$ are saved. To make use of the archive, we will consider that the individual $p_{r2,G}$ from Equation \eqref{eq3} is selected from $P \cup A$, which is the union of the population and the file $A$. The file size is the same as the population size. When the size of $A$ exceeds the size $P$, a random individual is selected to remove it and make room for the new one.

The parameters $p$, $F_i$ y $CR_i$ are difficult to adjust and their selection is not trivial, as they largely determine the success of the optimization process. These are the parameters that SHADE adaptively optimizes using the mentioned history record.

\subsection{SHADE Parameter Adaptation Method}

The SHADE method saves in memory a structure with $H$ entries for parameters $F_i$ and $CR_i$, as shown in Table \ref{tab:table2}. This structure is initialized with the value $0.5$ for all entries.

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{13pt}
	%\renewcommand{\arraystretch}{0.9}
	\resizebox{\textwidth}{!}{
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline
		 Index & 1 & 2 & $\cdots$ & $H - 1$ & $H$ \\ 
		 \hline
		 \hline
		 $M_{CR}$ & $M_{CR,1}$ & $M_{CR,1}$ & $\cdots$ & $M_{CR,H-1}$ & $M_{CR,H}$ \\ 
		 \hline
		 $M_{F}$ & $M_{F,1}$ & $M_{F,1}$ & $\cdots$ & $M_{F,H-1}$ & $M_{F,H}$ \\ 
		\hline

	\end{tabular}}
	\caption{The historical memory $M_{CR}$, $M_{F}$  \cite{tanabe2013success}}
	\label{tab:table2}
\end{table}

In each generation, parameters $CR_i$ and $F_i$ are calculated on the basis of the history by applying the Equations \eqref{eq6} and \eqref{eq7}:

\begin{equation}
CR_i = randn_i(M_{CR,r_i}, 0.1)
\label{eq6}
\end{equation}

\begin{equation}
F_i = randc_i(M_{F,r_i}, 0.1)
\label{eq7}
\end{equation}

Where $randn(\mu, \sigma^2)$ and $randc(\mu, \sigma^2)$ are random values from normal and Cauchy distributions respectively, with mean $\mu$ and variance $\sigma^2$. When a value for $CR_i$ outside of the range $[0,1]$ is generated, its replaced by the allowed value. If $F_i > 1$, then it is truncated to $1$, and if $F_i < 0$, then Equation \eqref{eq7} is applied as many times as needed to obtain a legal value.

SHADE uses two auxiliar sets $S_{CR}$ and $S_F$ to store the $CR_i$ and $F_i$ values that were successfully used to generate a trial vector that replaced the parent. At the end of each generation, the content of the memory $M$ is updated following Equations \eqref{eq8} and \eqref{eq9}.

\begin{equation}
M_{CR,h,G+1} = \left\{ \begin{array}{lc}
mean_{WA} (S_{CR}) &   if \;\; S_{CR} \neq \emptyset \\
M_{CR,h,G+1} &  otherwise
\end{array}
\right.
\label{eq8}
\end{equation}

\begin{equation}
M_{F,h,G+1} = \left\{ \begin{array}{lc}
mean_{WL} (S_{F}) &   if \;\; S_{F} \neq \emptyset \\
M_{F,h,G+1} &  otherwise
\end{array}
\right.
\label{eq9}
\end{equation}

The $h \;\; (0 \le h \le H)$ index specifies the position of the memory $M$ to be updated. This index is initialized to $0$ at the beginning of the optimization process and increased by one after each generation. When $h \ge H$ then $h$ is restarted to 1. It is worth noting that, when in a generation there is no trial vector better than the parent which generated it, no update of $M$ is done.

In Equation \ref{eq8}, the term $mean_{WA} (S_{CR})$ is the wighted mean,  which is computed following Equations \eqref{eq10} and \eqref{eq11}, proposed by Peng et al (2009) \cite{peng2009multi} to prevent $CR$ to converge to small values.

\begin{equation}
mean_{WA} (S_{CR}) = \sum_{i = 1}^{|S_{RC}|} \omega_i * S_{RC,i}
\label{eq10}
\end{equation}

\begin{equation}
\omega_i = \frac{\Delta f_i}{\sum_{i = 1}^{|S_{RC}|} \Delta f_i}
\label{eq11}
\end{equation}

where $\Delta f_i = |f(t_i,G) - f(p_i, G)|$, which is the amount of improvement obtained from the trial vector respect to to the parent.

In Equation \ref{eq9} the term $mean_{WL} (S_{F})$ refers to the weighted Lehmer mean, which is computed as in Equation \eqref{eq12}, proposed by Tanabe and Fukunaga (2013) \cite{tanabe2013success}.

\begin{equation}
mean_{WL} (S_{F}) = \frac{\sum_{i = 1}^{|S_{F}|} \omega_i * S^2_{F,i}}{\sum_{i = 1}^{|S_{F}|} \omega_i * S_{F,i}}
\label{eq12}
\end{equation}

Unlike parameters $S_{CR}$ and $S_F$, the parameter $pbest$ from Equation \eqref{eq3}, used to set the greedyness of the mutation strategy, is not included in the optimization process. However it is not static, it is calculated for each individual of the population $p_i$ following Equation \eqref{eq13}.

\begin{equation}
pbest_i = randn[p_{min}, 0.2]
\label{eq13}
\end{equation}

where $p_{min}$ is set such that there is always al least two individuals to choose between, so $p_{min} = 2/N$.

\subsection{Adaptation of SHADE for constrained clustering}

Para adaptar SHADE para resolver el porblema del clustering con restricciones tomaremos algunos de los elementos que permitian adaptarlo a BRKGA. Haremos uso del concepto de random-key para crear los individuos de la población, siendo cada individuo de la misma un vector de random-keys. Utilizaremos el mismo método de inicialización y decodificación que el empleado en BRKGA (consultar Sección \ref{sec:AdaptationofBRKGA}). Proponemos una nueva función fitness para evaluar los inidividuos de la población, definida en la Ecuación \eqref{eq14}, donde $z_i$ se obtiene como en la Ecuacion \eqref{eq2}.

To adapt SHADE to solve the constrained clustering problem we will take some of the elements that allowed us to adapt it to BRKGA. We will use random-keys to create the individuals of the population, this way each individual is defined as a vector of random-keys. We will also use the same initialization method and the same decoder used in BRKGA (see Section \ref{sec:AdaptationofBRKGA}).

For SHADE, we propose a new fitness function to evaluate the individuals of the population. Shade fitness function is defined in Equation \eqref{eq14}, where $z_i$ is obtained as in Equation \eqref{eq2}.

\begin{equation}
f_i = z_i * (infeasibility + 1)
\label{eq14}
\end{equation}

Note that in Equation \eqref{eq14} no parameter is involved that cannot be calculated based on the dataset or constraints, whereas in \eqref{eq1} the $\mu$ parameter must be calculated and optimized for each individual problem.

We found that with the fitness function defined in \eqref{eq1} there is no competition between the penalty term and the within-cluster-sum-of-squares term, because penalty is always way larger than it or zero. This can bias the  exploration of the solutions space to those that satisfies all the constraints, even if the within-cluster-sum-of-squares is still improvable by moving two instances implied in a constraint to different clusters without violating that constraint.

With Equation \eqref{eq14} we try to find a trade-off between the within-cluster-sum-of-squares and the penalty term, allowing solutions that violate a certain number of constraints to compete with those who violate a lesser number of them but find a better within-cluster-sum-of-squares.

\textcolor{red}{Poner una pintura que ejemplifique lo anterior.}\\
\textcolor{red}{Explicar la busqueda local. Con pseudocódigo?}

\clearpage

\section{Experimental Setup}

\subsection{Dataset}

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{0.9}
	\resizebox{\textwidth}{!}{
	\begin{tabular}{c c c c}
		\hline
		Name & Number of Instances & Number of Classes & Features \\
		\hline
		& & & \\
		& & & \\
		& & & \\
		\hline

	\end{tabular}}
	\caption{Summary of datasets used for the experiments \textcolor{red}{citar keel}}
	\label{undefined}
\end{table}

\subsection{Constraints Generation}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{0.9}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
	\begin{tabular}{lcc c cc c cc c cc}
		\hline
		\multirow{2}{*}{Dataset} &
		\multicolumn{2}{c}{Percent 1\%} &&  \multicolumn{2}{c}{Percent 5\%} && \multicolumn{2}{c}{Percent 15\%} && \multicolumn{2}{c}{Percent 20\%} \\
		\cline{2-3} \cline{5-6} \cline{8-9} \cline{11-12}
		& ML & CL && ML & CL && ML & CL && ML & CL \\
		\hline
		& & && & && & && & \\
		& & && & && & && & \\
		& & && & && & && & \\
		\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Number of constraints used in experiments}
	\label{undefined}
\end{table}

\subsection{Evaluation Method}

We use the evaluation method proposed by Wagstaff et al. (2001) \cite{wagstaff2001constrained}, the authors of the Constrained K-Means algorithm, proposed in their work to evaluate the results.

Since we have the true labels associated to each of the datasets, we can use them in post-processing to evaluate the results provided by each method. We will use Rand Index \cite{rand1971objective} to measure the accuracy of the predictions resulting from each method we test. Rand Index computes de degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$. $C_1$ and $C_2$ are viewed as collections of $N(N - 1)/2$ pairwise decisions.

For each pair of $x_i$ and $x_j$ instances in $X$, $C_i$ assigns them to the same cluster or to different clusters. We take $a$ as the number of pairings where $x_i$ is in the same cluster as $x_j$ in $C_1$ and $C_2$ , and $b$ as the opposite event. Then, the degree of similarity between $C_1$ and $C_2$ is calculated as in Equation \eqref{eq15}.

\begin{equation}
Rand(C_1, C_2) = \frac{a + b}{N(N - 1)/2}
\label{eq15}
\end{equation}

We will use this measure co compute the accuracy for all of our experiments.

\section{Experimental Results}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{0.9}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{l ccc c ccc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{3}{c}{BRKGA} &&  \multicolumn{3}{c}{SHADE} \\
			\cline{2-4} \cline{6-8}
			& RandIndex & Unsat(\%) & Time(s) && RandIndex & Unsat(\%) & Time(s) \\
			\hline
			&  &  &  &&  &  &  \\
			&  &  &  &&  &  &  \\
			&  &  &  &&  &  &  \\
			\hline
			Means &  &  &  &&  &  &  \\
			\hline

		\end{tabular}}
		%\end{adjustwidth}

	\caption{Experimental results obtained with X\% of labeled data}
	\label{undefined}
\end{table}

\section{Acknowledgement}

\section{Conclusions and Future Work}

\clearpage

\section*{References}

\bibliography{mybibfile}

\end{document}