### 2019-03-09 Papers/SHADE/descartados.tex
Correcciones en memoria SHADE y nuevos resultados

%%%%%%%%%%%%%%%%% Test frecuentiastas %%%%%%%%%%%%%%%%%%%

\subsection{Applying the Wilcoxon's signed-rank test to our results}

We will apply the Wilcoxon's signed-rank test for each of the measures obtained, namely: ARI, Unsat and Time. We set as null hypothesis $H_0: $ the distribution of $A - B$ is symmetric about $\mu$. However, given that in the case of ARI it is a problem of maximisation, and in the case of Time and Unsat it is a problem of minimisation, we will consider two different alternative hypotheses. In the case of maximization we propose $H_{max}:$ the distribution of $A - B$ is shifted to the left with respect to $\mu$. On the other hand, in the case of minimization we propose $H_{min}:$ the distribution of $A - B$ is shifted to the right with respect to $\mu$. In this way we assume that SHADE+LS is better than BRKGA+LS and we try to prove it.

In this case we set $\mu = 0$, and work with a confidence level of 95\%, which means that the significance level is 0.05 ($\alpha = 0.05$). Table \ref{tab:freqTest} shows the statistical values obtained using Wilcoxon's signed-rank test for each comparison.

\begin{table}[!htp]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	\begin{tabular}{ccccc}
		\hline
		Measure & $R^{+}$ & $R^{-}$ & P-value & Alternative hypothesis  \\
		\hline
		ARI & 3088 & 1665 & 0.005254 & $H_{max}$ \\
		Time & 4008 & 1042 & 0.0000001723 & $H_{min}$ \\
		Unsat & 3082.5 & 745.5 & 0.0000003838 & $H_{min}$ \\
		\hline
	\end{tabular}
	\caption{Results obtained by the Wilcoxon's signed-rank test for the comparison SHADE vs BRKGA}
	\label{tab:freqTest}
\end{table}

For all three considered measures, $P-value < \alpha$, therefore in all cases we can reject $H_0$ and infer that there are significant differences between the two methods, giving the advantage to SHADE+LS with basis to the proposed alternative hypothesis. In conclusion, we can safely admit that [the] Wilcoxon's signed-rank test provides empirical evidence in favour of SHADE+LS.