### 2019-03-23 Resultados/ResultadosTFG/Friedman/Friedman10/Friedman10.tex

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx}
\usepackage{lscape}
\title{Output tables for 1xN statistical comparisons.}
\author{}
\date{\today}
\begin{document}
\pagestyle{empty}
\maketitle
\thispagestyle{empty}

\section{Average rankings of Friedman test}

Average ranks obtained by each method in the Friedman test.

\begin{table}[!htp]
\centering
\begin{tabular}{|c|c|}\hline
Algorithm&Ranking\\\hline
"SHADE"&3.04\\"COPKM"&4.6\\"LCVQE"&2.7\\"RDPM"&2.78\\"TVClust"&3.22\\"CEKM"&4.66\\\hline\end{tabular}
\caption{Average Rankings of the algorithms (Friedman)}
\end{table}

Friedman statistic (distributed according to chi-square with 5 degrees of freedom): 28.6. \newline P-value computed by Friedman Test: 0.000028.\newline

\newpage

\section{Post hoc comparison (Friedman)}

P-values obtained in by applying post hoc methods over the results of Friedman procedure.

\begin{table}[!htp]
\centering\footnotesize
\begin{tabular}{cccccccc}
$i$&algorithm&$z=(R_0 - R_i)/SE$&$p$&Holm &Holland&Finner&Li\\
\hline5&"CEKM"&3.704052&0.000212&0.01&0.010206&0.010206&0.006325\\4&"COPKM"&3.590662&0.00033&0.0125&0.012741&0.020308&0.006325\\3&"TVClust"&0.982708&0.325751&0.016667&0.016952&0.030307&0.006325\\2&"SHADE"&0.64254&0.520523&0.025&0.025321&0.040204&0.006325\\1&"RDPM"&0.151186&0.879829&0.05&0.05&0.05&0.05\\\hline
\end{tabular}
\caption{Post Hoc comparison Table for $\alpha=0.05$ (FRIEDMAN)}
\end{table}Holm's procedure rejects those hypotheses that have an unadjusted p-value $\le0.016667$.

Holland's procedure rejects those hypotheses that have an unadjusted p-value $\le0.016952$.

Finner's procedure rejects those hypotheses that have an unadjusted p-value $\le0.030307$.

Li's procedure rejects those hypotheses that have an unadjusted p-value $\le0.006325$.

\newpage

\section{Adjusted P-Values (Friedman)}

Adjusted P-values obtained through the application of the post hoc methods (Friedman).

\begin{table}[!htp]
\centering\small
\begin{tabular}{cccc}
i&algorithm&unadjusted $p$&$p_{Holm}$\\
\hline1&"CEKM"&0.000212&0.001061\\2&"COPKM"&0.00033&0.001319\\3&"TVClust"&0.325751&0.977254\\4&"SHADE"&0.520523&1.041046\\5&"RDPM"&0.879829&1.041046\\\hline
\end{tabular}
\caption{Adjusted $p$-values (FRIEDMAN) (I)}
\end{table}
\begin{table}[!htp]
\centering\small
\begin{tabular}{cccccc}
i&algorithm&unadjusted $p$&$p_{Holland}$&$p_{Finner}$&$p_{Li}$\\
\hline1&"CEKM"&0.000212&0.00106&0.00106&0.001763\\2&"COPKM"&0.00033&0.001319&0.00106&0.002737\\3&"TVClust"&0.325751&0.693479&0.481558&0.730512\\4&"SHADE"&0.520523&0.770102&0.601012&0.812436\\5&"RDPM"&0.879829&0.879829&0.879829&0.879829\\\hline
\end{tabular}
\caption{Adjusted $p$-values (FRIEDMAN) (II)}
\end{table}

\newpage
\end{document}