### 2019-03-23 Resultados/ResultadosTFG/Friedman/Friedman20/Friedman20.tex

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx}
\usepackage{lscape}
\title{Output tables for 1xN statistical comparisons.}
\author{}
\date{\today}
\begin{document}
\pagestyle{empty}
\maketitle
\thispagestyle{empty}

\section{Average rankings of Friedman test}

Average ranks obtained by each method in the Friedman test.

\begin{table}[!htp]
\centering
\begin{tabular}{|c|c|}\hline
Algorithm&Ranking\\\hline
"SHADE"&2.8\\"COPKM"&2.86\\"LCVQE"&3.66\\"RDPM"&3.28\\"TVClust"&3.3\\"CEKM"&5.1\\\hline\end{tabular}
\caption{Average Rankings of the algorithms (Friedman)}
\end{table}

Friedman statistic (distributed according to chi-square with 5 degrees of freedom): 25.525714. \newline P-value computed by Friedman Test: 0.00011.\newline

\newpage

\section{Post hoc comparison (Friedman)}

P-values obtained in by applying post hoc methods over the results of Friedman procedure.

\begin{table}[!htp]
\centering\footnotesize
\begin{tabular}{cccccccc}
$i$&algorithm&$z=(R_0 - R_i)/SE$&$p$&Holm &Holland&Finner&Li\\
\hline5&"CEKM"&4.346591&0.000014&0.01&0.010206&0.010206&0.004751\\4&"LCVQE"&1.625247&0.10411&0.0125&0.012741&0.020308&0.004751\\3&"TVClust"&0.944911&0.344704&0.016667&0.016952&0.030307&0.004751\\2&"RDPM"&0.907115&0.364346&0.025&0.025321&0.040204&0.004751\\1&"COPKM"&0.113389&0.909722&0.05&0.05&0.05&0.05\\\hline
\end{tabular}
\caption{Post Hoc comparison Table for $\alpha=0.05$ (FRIEDMAN)}
\end{table}Holm's procedure rejects those hypotheses that have an unadjusted p-value $\le0.0125$.

Holland's procedure rejects those hypotheses that have an unadjusted p-value $\le0.012741$.

Finner's procedure rejects those hypotheses that have an unadjusted p-value $\le0.020308$.

Li's procedure rejects those hypotheses that have an unadjusted p-value $\le0.004751$.

\newpage

\section{Adjusted P-Values (Friedman)}

Adjusted P-values obtained through the application of the post hoc methods (Friedman).

\begin{table}[!htp]
\centering\small
\begin{tabular}{cccc}
i&algorithm&unadjusted $p$&$p_{Holm}$\\
\hline1&"CEKM"&0.000014&0.000069\\2&"LCVQE"&0.10411&0.41644\\3&"TVClust"&0.344704&1.034113\\4&"RDPM"&0.364346&1.034113\\5&"COPKM"&0.909722&1.034113\\\hline
\end{tabular}
\caption{Adjusted $p$-values (FRIEDMAN) (I)}
\end{table}
\begin{table}[!htp]
\centering\small
\begin{tabular}{cccccc}
i&algorithm&unadjusted $p$&$p_{Holland}$&$p_{Finner}$&$p_{Li}$\\
\hline1&"CEKM"&0.000014&0.000069&0.000069&0.000153\\2&"LCVQE"&0.10411&0.355803&0.240309&0.535578\\3&"TVClust"&0.344704&0.718608&0.505618&0.792456\\4&"RDPM"&0.364346&0.718608&0.505618&0.801423\\5&"COPKM"&0.909722&0.909722&0.909722&0.909722\\\hline
\end{tabular}
\caption{Adjusted $p$-values (FRIEDMAN) (II)}
\end{table}

\newpage
\end{document}