### 2019-06-26 Memoria/Chapters/ExperimentalResults.tex
Nuevos capitulos en memoria y correcciones en papers

\chapter{Experimental Results} \label{ch:ExperimentalResults}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.2}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{l ccc c ccc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
			\cline{2-4} \cline{6-8}
			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
			\hline
			Appendicitis & 0.022 & 5.454 & 383.642 && \textbf{0.086} & 5.091 & 366.491 \\
			Breast Cancer & \textbf{0.395} & 17.444 & 3651.263 && 0.159 & 22.256 & 3602.665 \\
			Bupa & \textbf{0.281} & 14.151 & 1420.117 && 0.104 & 16.672 & 1409.487 \\
			Circles & \textbf{0.166} & 12.598 & 1077.238 && 0.063 & 14.207 & 1060.787 \\
			Ecoli & 0.009 & 20.321 & 1169.798 && \textbf{0.020} & 12.763 & 1112.981 \\
			Glass & \textbf{0.016} & 8.658 & 735.507 && 0.013 & 4.848 & 694.962 \\
			Haberman & 0.047 & 15.699 & 1155.985 && \textbf{0.084} & 15.441 & 1146.023 \\
			Hayesroth & \textbf{0.027} & 1.500 & 550.117 && 0.025 & 2.000 & 529.424 \\
			Ionosphere & 0.125 & 17.587 & 1894.390 && \textbf{0.332} & 12.603 & 1872.308 \\
			Iris & 0.238 & 0.190 & 505.112 && \textbf{0.342} & 0.000 & 483.584 \\
			Monk2 & \textbf{0.421} & 14.926 & 1848.543 && 0.270 & 16.385 & 1821.208 \\
			Moons & \textbf{0.277} & 11.816 & 1098.089 && 0.272 & 11.356 & 1075.089 \\
			Saheart & 0.147 & 20.555 & 2135.624 && \textbf{0.265} & 16.725 & 2095.554 \\
			Sonar & \textbf{0.124} & 7.619 & 1106.039 && 0.116 & 8.952 & 1081.994 \\
			Soybean & 0.496 & 0.000 & 176.333 && \textbf{0.515} & 0.000 & 151.163 \\
			Spectfheart & 0.334 & 9.573 & 1451.235 && \textbf{0.343} & 8.604 & 1405.243 \\
			Spiral & 0.058 & 15.402 & 1098.622 && \textbf{0.114} & 13.793 & 1099.747 \\
			Tae & 0.018 & 0.833 & 520.855 && \textbf{0.024} & 1.333 & 508.988 \\
			Vowel & \textbf{0.003} & 13.330 & 4245.672 && 0.002 & 11.523 & 4290.154 \\
			Zoo & \textbf{0.105} & 1.091 & 336.345 && 0.098 & 2.182 & 317.674 \\
			\hline
			Mean & \textbf{0.161} & 11.771 & 1675.618 && 0.156 & 10.828 & 1652.02 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{10}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
	\label{tab:results10}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.2}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{l ccc c ccc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
			\cline{2-4} \cline{6-8}
			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
			\hline
			Appendicitis & 0.296 & 6.167 & 381.625 && \textbf{0.332} & 5.833 & 362.106 \\
			Breast Cancer & 0.772 & 10.172 & 3854.911 && \textbf{0.911} & 3.310 & 3766.402 \\
			Bupa & \textbf{0.840} & 5.973 & 1463.152 && 0.801 & 5.641 & 1437.639 \\
			Circles & 0.828 & 6.101 & 1136.484 && \textbf{0.982} & 0.404 & 1108.928 \\
			Ecoli & 0.015 & 24.314 & 1200.463 && \textbf{0.025} & 19.090 & 1127.568 \\
			Glass & 0.017 & 17.500 & 745.987 && \textbf{0.031} & 11.743 & 722.009 \\
			Haberman & \textbf{0.866} & 4.677 & 1209.471 && 0.796 & 5.314 & 1147.479 \\
			Hayesroth & \textbf{0.107} & 7.536 & 545.861 && 0.099 & 8.261 & 520.702 \\
			Ionosphere & \textbf{0.803} & 7.722 & 1986.543 && 0.591 & 11.974 & 1944.085 \\
			Iris & \textbf{0.412} & 3.873 & 506.548 && 0.260 & 6.087 & 486.723 \\
			Monk2 & 0.751 & 10.827 & 1949.305 && \textbf{0.945} & 2.067 & 1920.361 \\
			Moons & 0.958 & 1.334 & 1133.858 && \textbf{0.995} & 0.020 & 1113.687 \\
			Saheart & 0.749 & 10.940 & 2281.992 && \textbf{0.945} & 1.839 & 2225.210 \\
			Sonar & 0.800 & 5.040 & 1124.700 && \textbf{0.890} & 2.177 & 1069.862 \\
			Soybean & \textbf{0.428} & 0.000 & 170.674 && 0.422 & 0.000 & 155.948 \\
			Spectfheart & \textbf{0.924} & 2.854 & 1561.040 && 0.778 & 5.732 & 1485.260 \\
			Spiral & 0.847 & 5.677 & 1133.962 && \textbf{0.949} & 1.495 & 1102.920 \\
			Tae & \textbf{0.119} & 7.115 & 522.088 && 0.066 & 8.380 & 513.289 \\
			Vowel & \textbf{0.002} & 14.486 & 4807.165 && 0.001 & 13.341 & 4867.849 \\
			Zoo & \textbf{0.157} & 3.500 & 333.984 && 0.128 & 2.834 & 315.131 \\
			\hline
			Mean & 0.483 & 9.456 & 1784.931 && \textbf{0.507} & 6.997 & 1751.368 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{15}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
	\label{tab:results15}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.2}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{l ccc c ccc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
			\cline{2-4} \cline{6-8}
			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
			\hline
			Appendicitis & 1.000 & 0.000 & 389.905 && 1.000 & 0.000 & 371.579 \\
			Breast Cancer & 0.807 & 9.086 & 4061.162 && \textbf{0.977} & 0.891 & 3984.370 \\
			Bupa & 0.936 & 2.975 & 1529.151 && \textbf{0.993} & 0.281 & 1508.772 \\
			Circles & 0.935 & 2.814 & 1180.268 && \textbf{1.000} & 0.000 & 1170.452 \\
			Ecoli & 0.028 & 25.716 & 1267.419 && \textbf{0.047} & 21.466 & 1277.903 \\
			Glass & 0.038 & 23.610 & 766.861 && \textbf{0.082} & 16.412 & 796.162 \\
			Haberman & 0.943 & 2.073 & 1266.131 && \textbf{0.997} & 0.138 & 1235.441 \\
			Hayesroth & \textbf{0.548} & 6.331 & 554.339 && 0.531 & 7.056 & 540.585 \\
			Ionosphere & 0.902 & 4.314 & 2051.251 && \textbf{0.993} & 0.258 & 2019.558 \\
			Iris & \textbf{0.691} & 5.241 & 513.675 && 0.553 & 6.528 & 488.262 \\
			Monk2 & 0.855 & 6.025 & 2055.425 && \textbf{0.991} & 0.379 & 2013.468 \\
			Moons & 0.937 & 2.689 & 1177.848 && \textbf{0.995} & 0.158 & 1157.222 \\
			Saheart & 0.881 & 5.250 & 2402.513 && \textbf{0.981} & 0.706 & 2353.931 \\
			Sonar & 0.992 & 0.186 & 1138.769 && \textbf{1.000} & 0.000 & 1127.020 \\
			Soybean & 0.416 & 0.000 & 171.281 && \textbf{0.563} & 0.000 & 153.406 \\
			Spectfheart & 0.953 & 1.677 & 1602.229 && \textbf{1.000} & 0.000 & 1526.925 \\
			Spiral & 0.921 & 3.604 & 1178.580 && \textbf{0.998} & 0.091 & 1155.194 \\
			Tae & 0.351 & 9.807 & 527.292 && \textbf{0.392} & 9.419 & 508.610 \\
			Vowel & 0.002 & 14.682 & 5450.331 && \textbf{0.003} & 13.766 & 5514.052 \\
			Zoo & \textbf{0.176} & 5.429 & 335.715 && 0.153 & 4.191 & 320.186 \\
			\hline
			Mean & 0.602 & 8.086 & 1895.671 && \textbf{0.647} & 5.443 & 1880.921 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{20}$ comparing SHADE$_{CC}$ and BRKGA$_{CC}$.}
	\label{tab:results20}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & \textbf{0.611} & 0.022 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
			Breast Cancer & \textbf{0.755} & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
			Bupa & \textbf{0.870} & 0.281 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
			Circles & \textbf{0.798} & 0.166 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
			Ecoli & 0.069 & 0.009 & - & 0.387 & \textbf{0.417} & 0.265 & - \\
			Glass & 0.009 & 0.016 & 0.184 & \textbf{0.268} & 0.197 & 0.211 & - \\
			Haberman & \textbf{0.638} & 0.047 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
			Hayesroth & 0.031 & 0.027 & - & \textbf{0.106} & 0.097 & 0.054 & 0.139 \\
			Ionosphere & \textbf{0.792} & 0.125 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
			Iris & 0.621 & 0.238 & -0.285 & \textbf{0.730} & 0.607 & 0.244 & 0.684 \\
			Monk2 & 0.815 & 0.421 & \textbf{0.982} & 0.072 & 0.094 & -0.002 & 0.007 \\
			Moons & \textbf{0.968} & 0.277 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
			Saheart & 0.800 & 0.147 & \textbf{0.974} & 0.018 & 0.020 & 0.068 & 0.000 \\
			Sonar & \textbf{0.649} & 0.124 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
			Soybean & 0.304 & 0.496 & 0.503 & 0.545 & \textbf{0.621} & 0.000 & 0.244 \\
			Spectfheart & \textbf{0.896} & 0.334 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
			Spiral & \textbf{0.856} & 0.058 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
			Tae & \textbf{0.031} & 0.018 & - & 0.009 & -0.000 & 0.067 & 0.000 \\
			Vowel & 0.002 & 0.003 & - & 0.063 & -0.003 & \textbf{0.067} & - \\
			Zoo & 0.151 & 0.105 & \textbf{0.715} & 0.666 & 0.412 & 0.335 & - \\
			\hline
			Mean & \textbf{0.474} & 0.158 & -0.510 & 0.226 & 0.212 & 0.175 & -0.186 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{10}$ by DILS$_{CC}$ and the state-of-the-art methods.}
	\label{tab:results10}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & \textbf{0.957} & 0.296 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
			Breast Cancer & 0.788 & 0.772 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.037 \\
			Bupa & 0.984 & 0.840 & \textbf{1.000} & -0.005 & -0.007 & -0.007 & 0.001 \\
			Circles & \textbf{1.000} & 0.828 & \textbf{1.000} & -0.003 & 0.375 & 0.973 & 0.105 \\
			Ecoli & 0.134 & 0.015 & - & 0.387 & 0.372 & \textbf{0.719} & - \\
			Glass & 0.057 & 0.017 & - & \textbf{0.280} & 0.253 & 0.229 & - \\
			Haberman & \textbf{1.000} & 0.866 & \textbf{1.000} & -0.002 & 0.075 & 0.219 & -0.050 \\
			Hayesroth & \textbf{0.386} & 0.107 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
			Ionosphere & 0.970 & 0.803 & \textbf{1.000} & 0.178 & 0.212 & 0.000 & 0.057 \\
			Iris & \textbf{0.850} & 0.412 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
			Monk2 & 0.902 & 0.751 & \textbf{1.000} & 0.072 & 0.170 & -0.002 & 0.007 \\
			Moons & \textbf{1.000} & 0.958 & \textbf{1.000} & 0.241 & 0.436 & 0.987 & 0.095 \\
			Saheart & 0.862 & 0.749 & \textbf{1.000} & 0.020 & 0.037 & 0.221 & 0.000 \\
			Sonar & \textbf{0.981} & 0.800 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
			Soybean & 0.469 & 0.428 & 0.584 & 0.545 & \textbf{0.605} & 0.000 & 0.000 \\
			Spectfheart & \textbf{1.000} & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
			Spiral & \textbf{1.000} & 0.847 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
			Tae & \textbf{0.299} & 0.119 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.073} & - \\
			Zoo & 0.185 & 0.157 & 0.435 & \textbf{0.642} & 0.450 & 0.353 & - \\
			\hline
			Mean & \textbf{0.641} & 0.490 & -0.070 & 0.225 & 0.233 & 0.258 & -0.196 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{15}$ by DILS$_{CC}$ and the state-of-the-art methods.}
	\label{tab:results15}
\end{table}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & \textbf{1.000} & \textbf{1.000} & - & 0.305 & 0.331 & 0.012 & -0.006 \\
			Breast Cancer & 0.807 & 0.807 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
			Bupa & 0.991 & 0.936 & \textbf{1.000} & -0.005 & -0.007 & -0.006 & 0.000 \\
			Circles & \textbf{1.000} & 0.935 & \textbf{1.000} & -0.003 & 0.629 & 1.000 & 0.138 \\
			Ecoli & 0.201 & 0.028 & - & 0.387 & 0.459 & \textbf{0.763} & - \\
			Glass & \textbf{0.322} & 0.038 & - & 0.271 & 0.287 & 0.218 & - \\
			Haberman & \textbf{1.000} & 0.943 & \textbf{1.000} & -0.002 & 0.106 & 0.737 & 0.000 \\
			Hayesroth & \textbf{0.812} & 0.548 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
			Ionosphere & 0.982 & 0.902 & \textbf{1.000} & 0.173 & 0.309 & 0.000 & 0.115 \\
			Iris & \textbf{0.968} & 0.691 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
			Monk2 & 0.904 & 0.855 & \textbf{1.000} & 0.072 & 0.253 & -0.002 & 0.160 \\
			Moons & \textbf{1.000} & 0.937 & \textbf{1.000} & 0.241 & 0.831 & 1.000 & 0.180 \\
			Saheart & 0.881 & 0.881 & \textbf{1.000} & 0.018 & 0.026 & \textbf{1.000} & -0.006 \\
			Sonar & \textbf{1.000} & 0.992 & \textbf{1.000} & 0.004 & 0.127 & 0.000 & 0.003 \\
			Soybean & 0.564 & 0.416 & -0.218 & 0.545 & \textbf{0.631} & 0.000 & -0.016 \\
			Spectfheart & \textbf{1.000} & 0.953 & \textbf{1.000} & -0.107 & -0.112 & 0.000 & -0.054 \\
			Spiral & \textbf{1.000} & 0.921 & \textbf{1.000} & -0.003 & 0.011 & 0.006 & 0.045 \\
			Tae & \textbf{0.892} & 0.351 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.071} & - \\
			Zoo & 0.203 & 0.176 & \textbf{0.821} & 0.642 & 0.439 & 0.335 & - \\
			\hline
			Mean & \textbf{0.733} & 0.642 & 0.064 & 0.223 & 0.285 & 0.344 & -0.187 \\
			\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{20}$ by DILS$_{CC}$ and the state-of-the-art methods.}
	\label{tab:results20}
\end{table}