### 2019-06-26 Papers/DILS/paperDILS.tex

\documentclass[review]{elsarticle}

\usepackage{graphics, float, url}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{lineno,hyperref}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{chngpage}
\usepackage{setspace}
\usepackage{amsfonts}
\usepackage{subcaption}
\usepackage{float}
\usepackage[ruled, lined, onelanguage, linesnumbered]{algorithm2e}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\newcommand{\myfloatalign}{\centering}
\modulolinenumbers[5]

\usepackage{epstopdf}
\epstopdfDeclareGraphicsRule{.tiff}{png}{.png}{convert #1 \OutputFile}
\AppendGraphicsExtensions{.tiff}

\journal{Computers and Operations Research}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%
\newtheorem{definition}{Definition}
\begin{document}

\begin{frontmatter}

\title{Improving instance-level constrained clustering through Dual Iterative Local Search}

\author[mymainaddress]{Germ\'an Gonz\'alez-Almagro\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{germangalmagro@ugr.es}

\author[mymainaddress]{Juli\'an Luengo}

\author[mysecondaddress]{Jos\'e-Ram\'on Cano}

\author[mymainaddress]{Salvador Garc\'ia}

\address[mymainaddress]{DaSCI Andalusian Institute of Data Science and Computational Intelligence, University of Granada, Spain}

\address[mysecondaddress]{Dept. of Computer Science, EPS of Linares, University of Ja\'en, Campus Cient\'ifico Tecnol\'ogico de Linares, Cintur\'on Sur S/N, Linares 23700, Ja\'en, Spain}

\begin{abstract}
Clustering has always been a powerful tool in knowledge discovery. Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning: constrained clustering. This technique is a generalization of traditional clustering that considers additional information encoded by constraints. Constraints can be given in the form of instance-level
must-link and cannot-link constraints, which this paper focuses on. We propose a new metaheuristic algorithm, the Dual Iterative Local Search, and prove its capability to produce quality results for the constrained clustering problem. We compare the results obtained by this proposal to those obtained by previous nature-inspired techniques and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
\end{abstract}

\begin{keyword}
constrained clustering, instance-level, must-link, cannot-link, dual iterative local search.
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction} \label{sec:Intro}

One of the most widely known and studied data analysis problems is clustering. It is one of the most successful methods in the field of unsupervised learning, where there is no supervision on how the information should be handled. Clustering has been able to provide solutions in a large number of knowledge fields, such as, marketing, banking, psychology, psychiatry, astronomy, archaeology, genetics,\cite{Everitt:2009:CA:1538772} cartographic labeling \cite{araujo2019improving} and image segmentation in various fields, ranging from general applications \cite{wang2018non} to very specific ones such as in medicine \cite{verma2016improved, aparajeeta2016modified}.

In the literature, clustering methods are often divided into two subsets: partitional clustering and hierarchical clustering. In hierarchical clustering the result is not a partition of the data with a certain number of clusters---as in partitional clustering---, but instead a dendrogram in which there are partitions that include from the whole dataset to particular individuals \cite{Everitt:2009:CA:1538772}. A representative example of partitional clustering is the widely studied and well-known algorithm K-means \cite{wu2009top}, while for hierarchical clustering the CURE method should be highlighted \cite{guha1998cure}. In this paper we will focus on partitional clustering.

We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them. A dataset $X$ is composed of $n$ instances, each one of them described by $u$ features. More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,u]})$. A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$. As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$. The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement \cite{jain1999data}.

Semi-supervised learning (SSL) is a machine learning paradigm that arises from adding incomplete information to unsupervised learning. We can divide SSL methods into two broad categories according to their objective: semi-supervised classification and semi-supervised clustering \cite{chapelle2009semi}. The first method has partial information about the labels, so it tries to minimize the error based on them while taking into account the distribution of unlabeled instances. The second tries to obtain better defined clusters incorporating background information to the clustering process, this is known as constrained clustering and is the main subject of the study presented in this paper \cite{triguero2015self}.

Constrained clustering is an SSL learning method whose goal is to find a partition of the dataset that meets the proper characteristics of a clustering method result, in addition to satisfying a certain constraint set. It has been successfully applied in many knowledge fields, among which it is worth mentioning: advanced robotics applications \cite{davidson2005clustering, semnani2016constrained}, applied marketing \cite{seret2014new}, obstructive sleep apnea analysis \cite{mai2018evolutionary}, handwritten digits classification \cite{li2015scalable}, Internet traffic classification \cite{wang2014internet}, electoral district designing, \cite{brieden2017constrained}, and lane finding in GPS data \cite{wagstaff2001constrained} among others.

Constraints can be understood in different ways, resulting in three main types of constrained clustering: cluster-level \cite{bradley2000constrained}, instance-level \cite{davidson2007survey} and feature-level constrained clustering \cite{schmidt2011clustering}. Moreover, hybrid approaches which try to integrate different types of constraints have also been proposed \cite{wang2010clustering}.

In particular, we can find in the literature two types of instance-level constraints: pairwise constraints and distance-based constraints. In one hand, pairwise constraints tell us if two specific instances of a dataset must be placed in the same or in different clusters, resulting in Must-link (ML) and Cannot-link (CL) constraints respectively. On the other hand, distance-based constraints do not involve specific instances, but tell us if instances must be placed in the same or in different clusters based on a given distance measure \cite{davidson2007survey}. This paper focuses on pairwise instance-level constraints (ML and CL), which will be discussed later in Section \ref{sec:BackCC}.

Regarding the degree to which the constraints have to be satisfied, we can make a distinction between the concepts of hard \cite{wagstaff2001constrained} and soft \cite{law2004clustering} constraints. Hard constraints must necessarily be satisfied in the output partition of any algorithm that makes use of them, while soft constraints are taken as a strong guide for the algorithm that uses them but can be partially satisfied in the output partition \cite{seret2014new}. For the purposes of this paper, we will employ the latter.

Finding the optimal partition in a dataset, with respect to any kind of reasonable criteria, is known to be a $\mathbf{NP}$-hard problem. Therefore, the incorporation of constraints may modify the complexity of the clustering problem, depending on the type of constraints used. As we will study in more depth in Section \ref{sec:BackFeas}, the use of ML and CL constraints makes the constrained clustering problem $\mathbf{NP}$-complete \cite{davidson2005clustering}.

The constrained clustering problem can be formulated in terms of optimization, so that we can apply various optimization techniques to solve it. As mentioned earlier, its resolution poses a tough challenge, and metaheuristics are presented as a promising option to find quality approximate solutions. Metaheuristic algorithms are designed to explore the solution space of a problem with the guidance of a fitness (heuristic) function \cite{Gendreau:2010:HM:1941310}. In this type of approaches the key is to find a good exploration-exploitation tradeoff. This is the reason why the classic Local Search (LS) heuristic algorithm derived in the Iterative Local Search (ILS) metaheuristic algorithm, which introduces periodic perturbations in the exploration of the solution space to escape local optima \cite{lourencco2010iterated}. ILS---or variants of it---has been applied to a wide variety of problems, including the traveling salesman \cite{archetti2018iterated}, the quadratic multiple knapsack \cite{avci2017multi} or the vehicle routing problem \cite{chentli2018impact, estrada2019biased}. ILS is also very commonly used in the design of hybrid methods with good exploration-exploitation tradeoff, such as its combination with expediting mechanisms \cite{zohali2019reformulation} or with the success history based differential evolution algorithm \cite{zhao2019hybrid}.

In this paper we propose a new variant of the ILS method, the Dual Iterative Local Search (DILS), which introduces diversity in the exploration of the solution space in an adaptive and guided way. We apply it to the constrained clustering problem, proving that it constitutes a competitive approach to obtain quality results.

Regarding the organization of this paper, Section \ref{sec:background} reviews the existing knowledge concerning constrained clustering and the state-of-the-art. In Sections \ref{sec:DILS} and \ref{sec:DILS_CC} the DILS algorithm and its formulation for constrained clustering will be reviewed. Sections \ref{sec:expSetup} to \ref{sec:analisis} present the experimental setup, the results and their analysis, respectively. Finally, in Section \ref{sec:conclusiones} the conclusions are discussed.

\section{Background} \label{sec:background}

In this section we present the background knowledge concerning constrained clustering (Section \ref{sec:BackCC}), its computational complexity (Section \ref{sec:BackFeas}) and a brief description of some of the state-of-the-art methods for constrained clustering (Section \ref{sec:BackSOTA}).

\subsection{Constrained Clustering} \label{sec:BackCC}

In most clustering applications it is common to have some kind of information about the dataset to be analyzed. In pairwise instance-level constrained clustering this information is given in the form of pairs of instances. A constraint states whether the instances which it refers to must, or must not, be assigned to the same cluster. It is possible to obtain a better result by using this type of information than by using completely unsupervised clustering algorithms. We can now formalize the two type of constraints mentioned:

\begin{itemize}

	\item Must-link constraints $C_=(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ must be placed in the same cluster.

	\item Cannot-link constraints $C_{\neq}(x_i,x_j)$: instances $x_i$ and $x_j$ from $X$ cannot be assigned to the same cluster.

\end{itemize}

The goal of constrained clustering is to find a partition (or clustering) of $k$ clusters $C = \{c_1, \cdots, c_k\}$ of the dataset $X$ that ideally satisfies all constraints in the constraint set. As in the original clustering problem, it must be fulfilled that the sum of instances in each cluster $c_i$ is equal to the number of instances in $X$, which we have defined as $n = |X| = \sum_{i = 1}^{k} |c_i|$.

Knowing how a constraint is defined, ML constraints are an example of an equivalence relation; therefore, ML constraints are reflexive, transitive and symmetric. This way, given constraints $C_=(x_a,x_b)$ and $C_=(x_b,x_c)$ then $C_=(x_a,x_c)$ is verified. In addition, if $x_a \in c_i$ and $x_b \in c_j$ are related by $C_=(x_a,x_b)$, then $C_=(x_c,x_d)$ is verified for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.

It can also be proven that CL constraints do not constitute an equivalence relation. However, analogously, given $x_a \in c_i$ and $x_b \in c_j$, and the constraint $C_{\neq}(x_a,x_b)$, then it is also true that $C_{\neq}(x_c,x_d)$ for any $x_c \in c_i$ and $x_d \in c_j$ \cite{davidson2007survey}.

\subsection{The Feasibility Problem} \label{sec:BackFeas}

Given that constrained clustering adds a new element to the clustering problem, we must consider how this element affects the complexity of the problem. The feasibility problem for non-hierarchical instance-level constrained clustering was defined in \cite{davidson2005clustering} as in Definition \ref{def1}.

\begin{definition}

	\textbf{Feasibility Problem}: given a dataset $X$, a constraint set $CS$, and the bounds on the number of clusters $k_l \leq k \leq k_u$, does there exist a partition $C$ of $X$ with $k$ clusters such that all constraints in $CS$ are satisfied? \cite{davidson2005clustering}
	\label{def1}

\end{definition}

In \cite{davidson2005clustering} it is proven that, when $k_l = 1$ and $k_u \ge 3$, the feasibility problem for constrained clustering is $\mathbf{NP}$-complete, by reducing it from the Graph K-Colorability problem (it is also proven that it is not harder, so both have the same complexity). Table \ref{tab:feasibility} shows the complexity of the feasibility for different types of constraints.

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.2}
	%\resizebox{\textwidth}{!}{
		\begin{tabular}{c c}
			\hline
			Constraints & Complexity \\
			\hline
			Must-Link & $\mathbf{P}$\\
			Cannot-Link & $\mathbf{NP}$-complete\\
			ML and CL & $\mathbf{NP}$-complete\\
			\hline

		\end{tabular}%}
	\caption{Feasibility problem complexity \cite{davidson2005clustering}.}
	\label{tab:feasibility}
\end{table}

These complexity results show that the feasibility problem with CL constraints is intractable and hence constrained clustering is intractable too. For more details on the complexity of constrained clustering see \cite{davidson2005clustering}.

Intractable problems are hard to solve with deterministic and exact methods. That is the reason why metaheuristic algorithms constitute good approaches to find quality solutions to the constrained clustering problem.

\subsection{State-of-the-art Methods} \label{sec:BackSOTA}

The first adaptation of a classic clustering method for constrained clustering was proposed in \cite{wagstaff2001constrained}. It involved modifying the widely studied K-means algorithm to take into account instance-level constraints: the already known ML and CL. This method, named COP-kmeans, introduces a modification to the assignment rule of instances to clusters of the K-means algorithm so that an instance can be assigned to a cluster only if the assignment does not violate any constraint.

Within the fuzzy clustering family of methods, Constrained Evidential c-means (CECM), a variant of the Evidential c-means (ECM \cite{masson2008ecm}) algorithm, is proposed in \cite{antoine2012cecm}. The particularity of this algorithm is that the membership of instances to a cluster is defined by a probabilistic belief function. This method redefines constraints from the point of view of belief functions and includes them in the cost function.

A modification of the Constrained Vector Quantization Error algorithm (CVQE \cite{davidson2005clustering}) is proposed in \cite{pelleg2007k}. The CVQE algorithm proved to produce high quality results, at the cost of a very high computational complexity. Linear CVQE (LCVQE) introduces a modification of the cost function of CVQE to make it more intuitive and less computationally complex. The experimentation resulted in a dramatic improvement of clustering quality over both noisy and clean constraint sets.

Two Views Clustering (TVClust) and Relation Dirichlet Process - Means (RDPM) were proposed in \cite{khashabi2015clustering}. TVClust is able to incorporate the constraints into the clustering problem by making a soft interpretation of them. The authors model the dataset and constraints in different ways, perform clustering methods on them and try to find a consensus between both interpretations. Using this model as a basis, the authors derive the deterministic algorithm RDP-means. This method can be viewed as an extension of K-means that includes side information (constraints) and has the property that the number of clusters ($k$) does not need to be specified.

An adaptation of the Biased Random-Key Genetic Algorithm (BRKGA) \cite{gonccalves2011biased}---a variant of the Random-Key genetic algorithm \cite{bean1994genetic}---for the constrained clustering problem is proposed in \cite{de2017comparison}. It uses a genetic algorithm, combined with an LS procedure and guided by a penalty-style fitness function, to obtain a partition of the input dataset. It constitutes one of the few approaches to constrained clustering from the point of view of evolutionary computing.

\clearpage

\section{The Dual Iterative Local Search Method} \label{sec:DILS}

The Dual Iterative Local Search (DILS) is a new variant of the classic ILS method. Its goal is to perform a search in the solution space to find the solution with the best fitness value (given by fitness function $f$) by introducing diversity in an adaptive and guided way. While ILS works with a single individual, DILS keeps two of them $\{m_b, m_w\}$ in memory at all times, which allows it to guide diversity-inducing methods and to avoid local optima. The individual $m_b$ is the one providing the best fitness value, whereas $m_w$ provides the worst fitness value.

In each generation $G$ of the optimization process DILS builds a new individual by applying a recombination operator to $m_b$ and $m_w$. After that, it applies a strong mutation operator to the newly generated individual. This is the individual DILS applies the LS procedure to in order to improve its fitness value, resulting in the trial individual $m_t$. The trial individual $m_t$ replaces the worst of its parents, $m_w$, if it achieves a better fitness value; this operation represents the acceptance criterion for DILS. Equation \ref{eq1} displays the expression for this criterion in the case of a minimization problem.

\begin{equation}
m_{w,G} = \left\{ \begin{array}{lc}
m_t &   \text{if} \;\; f(m_t) < f(m_w)\\
m_{w,G} &  \text{otherwise}
\end{array}
\right..
\label{eq1}
\end{equation}

Now, the process described so far will most likely fall into a local optimum, effectively stopping the exploration of the solution space. To avoid it DILS implements a reinitialization method for $m_w$ based on the difference between the two individuals it keeps in memory ($\{m_b, m_w\}$). Equation \ref{eq2} shows the reinitialization criterion for a minimization problem.

\begin{equation}
m_{w,G+1} = \left\{ \begin{array}{lc}
\text{RandInit}() &   \text{if} \;\; f(m_b) - f(m_w) > f(m_b) * \xi \;\;\\
m_{w,G} &  \text{otherwise}
\end{array}
\right.,
\label{eq2}
\end{equation}

\noindent where RandInit$()$ is a function that returns a randomly initialized individual and $\xi \in [-1,1]$ is the parameter that controls the tolerance of the reinitialization method. This way the best individual is always preserved and therefore always takes part in the process of generating new individuals (via the recombination operator) as the best parent. Figure \ref{img:DILS_diag} summarizes the DILS optimization process.

\begin{figure}[!h]
	\centering
	\includegraphics[scale=0.25]{Figures/DILSDiagram.jpg}
	\caption{Diagram summarizing DILS optimization process.}\label{img:DILS_diag}
\end{figure}

It is crucial to note that in Figure \ref{img:DILS_diag} arrows in red indicate stages of the algorithm where $m_b$ and $m_w$ are reassigned to the true best and worst individual respectively. This way, at the point where the reinitialization criterion is tested, it is possible for $f(m_w)$ to yield a smaller value than $f(m_b)$, so that $f(m_b) - f(m_w)$ can be either positive or negative.

This highlights the influence of the parameter $\xi$. Let us assume again a minimization problem. On one hand, if $\xi > 0$, then the true worst individual is only reinitialized when $m_t$ replaces $m_w$ and is also better than $m_b$ by a margin. On the other hand, if $\xi < 0$, we allow DILS to reinitialize the true worst individual even when $m_t$ is not better than $m_b$. It becomes clear that when $\xi = 0$ the worst individual is restarted if $f(m_b) = f(m_w)$. Note that $\xi$ could take any real value, but it is not realistic to set it outside of the range $[-1,1]$, and we recommend $[-0.5, 0.5]$. To summarize, the closer $\xi$ is to 1 the more restrictive the reinitialization criterion is---in the minimization case. It is also worth noting that in early stages of the optimization process the reinitialization criterion tends to be met more often than in later stages. The reason for this is that the fitness value of the best individual $f(m_b)$ never worsens ($f(m_{b,G}) \ge f(m_{b,G+1})$), so it becomes increasingly harder for $m_w$ to achieve similar or better scores. Algorithm \ref{alg:DILS} summarizes the overall DILS optimization process.

\begin{algorithm}
	\SetNlSty{textbf}{[}{]}
	\SetNlSkip{0.5em}
	\setstretch{1.2}
	\SetKwFunction{RandInit}{RandInit}
	\SetKwFunction{Best}{Best}
	\SetKwFunction{Worst}{Worst}
	\SetKwFunction{Recombination}{Recombination}
	\SetKwFunction{Mutation}{Mutation}
	\SetKwFunction{LocalSearch}{LocalSearch}
	\KwIn{Probability for recombination operator $p_r$, segment size for mutation operator $p_s$, reinitialization method tolerance $\xi$.}
	\tcp{Initialization phase}
	$G \leftarrow 0$\\
	$m_b \leftarrow$ \RandInit{};
	$m_w \leftarrow$ \RandInit{}\\
	\tcp{Main loop}
	\While{Termination criteria are not met}{
		\tcp{Find best and worst individual}
		$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $};
		$m_w \leftarrow $ \Worst{$ \{m_b, m_w\} $}\\
		\tcp{Generate new individual}
		$m_t \leftarrow $ \Recombination{$m_b, \; m_w$}\\
		$m_t \leftarrow $ \Mutation{$m_t$}\\
		\tcp{Improve new individual}
		$m_t \leftarrow $ \LocalSearch{$m_t$}\\
		\tcp{Apply replacement operator}
		\If{$f(m_t) < f(m_w)$}{

			$m_w \leftarrow m_t$\\

		}
		\tcp{Check restart criterion}
		\If{$f(m_b) - f(m_w) > f(m_b) \times \xi$}{

			$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $}\\
			$m_w \leftarrow$ \RandInit{}\\

		}

		$G \leftarrow G + 1$
	}

	$m_b \leftarrow $ \Best{$ \{m_b, m_w\} $}\\
	\KwRet{$m_b$}
\caption{DILS}\label{alg:DILS}
\end{algorithm}

For the recombination and mutation operators that appear in Algorithm \ref{alg:DILS}, we use the uniform recombination and the segment mutation operator respectively. The uniform recombination consists in selecting features from the best individual $m_b$ based on a given probability $p_r$ to introduce them in the resulting individual, so the rest of the features are taken from $m_w$ \cite{spears1995virtues}. The segment mutation operator consists in replacing a fixed-size segment from the features of the individual with randomly generated features. The size of the segment $p_s$ must be fixed at the beginning of the optimization process.

It should be noted that the optimization process described so far is applicable to any minimization problem, although it is clear that there is a direct adaptation of DILS for maximization problems. In addition, the LS procedure used to obtain the trial individual $m_t$ must be specified for each problem, as well as numerical representation details and termination criteria.

\section{DILS optimization scheme for constrained clustering} \label{sec:DILS_CC}

In this section, we present an optimization scheme for the DILS algorithm when applied to constrained clustering. We will call this new approach DILS$_{CC}$. We will use an integer-based representation for all the individuals DILS$_{CC}$ work with; this way, each individual $m$ is defined as a vector of $n$ integers---with $n$ being the number of instances of the dataset. Each integer $m_i \in [0,k)$ in $m$ represents the label of the cluster that instance $x_i$ is assigned to, and therefore each individual $m$ is a solution to the clustering problem. Additionally, in order not to bias the exploration of the solution space, we will use random initialization to set $m_b$ and $m_w$; this procedure will also be used in the reinitialization of $m_w$.

Regarding the fitness function $f$, for each individual $m$ its fitness value $f_m$ can be calculated as in Equation \ref{eq3}.

\begin{equation}
f_m = z_m * (\text{infeasibility}_m + 1),
\label{eq3}
\end{equation}

\noindent where $\text{infeasibility}_m$ is the number of non-satisfied constraints and $z_m$ is the within-cluster-sum-of-squares which can be computed as in Equation \eqref{eq4}.

\begin{equation}
z_m = \sum_{c_i \in C_m} \left[ \frac{\sum_{x_a, x_b \in c_i} d^2(x_a,x_b)}{|c_i|}\right].
\label{eq4}
\end{equation}

We also need to specify an LS procedure for DILS to be able to improve the mutant individual to obtain the trial vector $m_t$. Algorithm \ref{alg:LS} shows the LS used in DILS$_{CC}$. Note that, in order not to bias the search, indices from vector $m$ and labels from the set of labels are randomly chosen to generate a random subset of the neighborhood of $m$. We also establish the maximum number of neighbors $p_m$ that can be generated in each call; this is done to ensure a better exploration-exploitation tradeoff.

\begin{algorithm}
	\SetNlSty{textbf}{[}{]}
	\SetNlSkip{0.5em}
	\setstretch{1.2}
	\SetKwFunction{RandomShuffle}{RandomShuffle}
	\SetKwFunction{Rand}{Rand}
	\SetKwRepeat{Do}{do}{while}
	\KwIn{Dataset $X$, constraint sets $C_=$ and $C_{\neq}$, individual (solution) $m$, number of clusters $k$, maximum number of neighbors that can be generated $p_m$.}
	\BlankLine
	\While{$improvement$ {\bf and} $generated < p_m$}{
		$improvement \leftarrow$ \texttt{false} \\
		\tcp{Select random index (object) from $m$}
		$i \leftarrow $ \Rand{$1$,$n$} \\
		\tcp{Random shuffle labels set}
		$RSL \leftarrow $ \RandomShuffle{$\{1,\cdots,K\}$}\\
		\For{$l \in RSL$}{
			$m^\prime \leftarrow m$\\
			\tcc{Move object $i$ from $m$ to the cluster associated with label $l$}
			$m^\prime_i \leftarrow l$\\
			\If{$f(m^\prime) < f(m)$}{
				$m \leftarrow m^\prime$\\
				$improvement \leftarrow$ \texttt{true} \\
			}
			$generated \leftarrow generated + 1$\\
		}
	}
	\BlankLine
	\KwRet{$m$}

	\caption{Local Search}\label{alg:LS}
\end{algorithm}

\clearpage

\section{Experimental Setup} \label{sec:expSetup}

For our experiments we will compare the results obtained by DILS$_{CC}$ and state-of-the-art methods over 25 datasets and 3 constraint sets for each one of them. Most of these datasets can be found at the \href{https://sci2s.ugr.es/keel/category.php?cat=clas}{Keel-dataset repository} \cite{triguero2017keel}, though some of them have been obtained via
\href{https://scikit-learn.org/stable/datasets/index.html}{\texttt{scikit-learn} python package} \cite{scikit-learn}. We also include 3 artificial datasets in our analysis, namely: \textit{Circles}, \textit{Moons} and \textit{Spiral}, which can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}. Table \ref{tab:datasets} displays a summary of every dataset.

\begin{table}[!h]
	\centering
	%\setlength{\arrayrulewidth}{1mm}
	%\setlength{\tabcolsep}{5pt}
	%\renewcommand{\arraystretch}{1.2}
	%\resizebox{\textwidth}{!}{
	\small
	\begin{tabular}{l c c c}
		\hline
		Name & No. Instances & No. Classes & No. Features \\
		\hline
		Appendicitis & 106 & 2 & 7 \\
		Breast Cancer & 569 & 2 & 30 \\
		Bupa & 345 & 2 & 6 \\
		Circles & 300 & 2 & 2 \\
		Ecoli & 336 & 8 & 7 \\
		Glass & 214 & 6 & 9 \\
		Haberman & 306 & 2 & 3 \\
		Hayesroth & 160 & 3 & 4 \\
		Heart & 270 & 2 & 13 \\
		Ionosphere & 351 & 2 & 33 \\
		Iris & 150 & 3 & 4 \\
		Led7Digit & 500 & 10 & 7 \\
		Monk2 & 432 & 2 & 6 \\
		Moons & 300 & 2 & 2 \\
		Movement Libras & 360 & 15 & 90 \\
		Newthyroid & 215 & 3 & 5 \\
		Saheart & 462 & 2 & 9 \\
		Sonar & 208 & 2 & 60 \\
		Soybean & 47 & 4 & 35 \\
		Spectfheart & 267 & 2 & 44 \\
		Spiral & 300 & 2 & 2 \\
		Tae & 151 & 3 & 5 \\
		Vowel & 990 & 11 & 13 \\
		Wine & 178 & 3 & 13 \\
		Zoo & 101 & 7 & 16 \\
		\hline

	\end{tabular}%}
	\caption{Summary of datasets used for the experiments.}
	\label{tab:datasets}
\end{table}

Classification datasets are commonly used in the literature to test constrained clustering algorithms; the reason behind this is that they enable us to generate constraints with respect to the true labels (see Section \ref{sec:ConstGent}). They also facilitate an easy evaluation of the quality of the algorithm by means of measures like the Adjusted Rand Index (see Section \ref{sec:EvalMet}).

\subsection{Constraint Generation} \label{sec:ConstGent}

Since we have the true labels associated with each dataset, we will use the method proposed in \cite{wagstaff2001constrained} to generate artificial constraint sets. This method consists of randomly selecting two instances of a dataset, then comparing its labels, and finally setting an ML or CL constraint depending on whether the labels are the same or different.

We will generate, for each dataset, three different sets of constraints---$CS_{10}$, $CS_{15}$ and $CS_{20}$---that will be associated with three small percentages of the size of the dataset: 10\%, 15\% and 20\%. With $n_f$ being the fraction of the size of the dataset associated with each of these percentages, the formula $\frac{n_f(n_f-1)}{2}$ tells us how many artificial constraints will be created for each constraint set; this number is equivalent to how many edges a complete graph with $n_f$ vertices would have.

The random allocation of constraints has a potential advantage over simply using the constraints contained in an $n_f$-vertex complete graph: there is a lower probability of biasing the constraint set towards having classes with poor representation in it. Table \ref{tab:constraints} shows the number of constraints of each type obtained for each dataset.

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.2}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
	\begin{tabular}{lcc c cc c cc}
		\hline
		\multirow{2}{*}{Dataset} &
		\multicolumn{2}{c}{$CS_{10}$} && \multicolumn{2}{c}{$CS_{15}$} && \multicolumn{2}{c}{$CS_{20}$} \\
		\cline{2-3} \cline{5-6} \cline{8-9}
		& ML & CL && ML & CL && ML & CL \\
		\hline
		Appendicitis & 39 & 16 && 71 & 49 && 164 & 67 \\
		Breast Cancer & 876 & 720 && 1965 & 1690 && 3487 & 2954 \\
		Bupa & 323 & 272 && 699 & 627 && 1201 & 1145 \\
		Circles & 208 & 227 && 502 & 488 && 853 & 917 \\
		Ecoli & 163 & 398 && 357 & 918 && 609 & 1669 \\
		Glass & 52 & 179 && 139 & 389 && 259 & 644 \\
		Haberman & 304 & 161 && 634 & 401 && 1135 & 756 \\
		Hayesroth & 39 & 81 && 102 & 174 && 177 & 319 \\
		Heart & 178 & 173 && 396 & 424 && 744 & 687 \\
		Ionosphere & 330 & 300 && 732 & 646 && 1299 & 1186 \\
		Iris & 26 & 79 && 82 & 171 && 136 & 299 \\
		Led7Digit & 126 & 1099 && 267 & 2508 && 460 & 4490 \\
		Monk2 & 473 & 473 && 979 & 1101 && 1917 & 1824 \\
		Moons & 200 & 235 && 494 & 496 && 900 & 870 \\
		Movement Libras & 27 & 603 && 112 & 1319 && 158 & 2398 \\
		Newthyroid & 108 & 123 && 270 & 258 && 449 & 454 \\
		Saheart & 595 & 486 && 1292 & 1123 && 2330 & 1948 \\
		Sonar & 100 & 110 && 245 & 251 && 436 & 425 \\
		Soybean & 4 & 6 && 6 & 22 && 12 & 33 \\
		Spectfheart & 233 & 118 && 543 & 277 && 965 & 466 \\
		Spiral & 224 & 211 && 487 & 503 && 918 & 852 \\
		Tae & 40 & 80 && 82 & 171 && 151 & 314 \\
		Vowel & 445 & 4406 && 1026 & 10000 && 1705 & 17798 \\
		Wine & 49 & 104 && 121 & 230 && 217 & 413 \\
		Zoo & 21 & 34 && 29 & 91 && 41 & 169 \\
		\hline

	\end{tabular}}
	%\end{adjustwidth}

	\caption{Number of constraints used in experiments.}
	\label{tab:constraints}
\end{table}

Note that the greater the number of classes present in the dataset, the fewer ML constraints obtained with the method proposed in \cite{wagstaff2001constrained}. This is because the probability of randomly choosing two individuals from the same class decreases as the number of classes present in the dataset increases.

\clearpage

\subsection{Evaluation Method} \label{sec:EvalMet}

Since we have the true labels associated to each of the datasets, we can use them to evaluate the results provided by each method. We will use the Adjusted Rand Index (ARI) to measure the accuracy of the predictions resulting from each method we test \cite{hubert1985comparing}. The basic Rand Index computes the degree of agreement between two partitions $C_1$ and $C_2$ of a given dataset $X$. $C_1$ and $C_2$ are viewed as collections of $n(n - 1)/2$ pairwise decisions \cite{rand1971objective}.

For each pair of instances $x_i$ and $x_j$ in $X$, a partition assigns them to the same cluster or to different clusters. We take $a$ as the number of pairings where $x_i$ is in the same cluster as $x_j$ in both $C_1$ and $C_2$, and $b$ as the opposite event ($x_i$ and $x_j$ are in different clusters in $C_1$ and $C_2$). Then, the degree of similarity between $C_1$ and $C_2$ is calculated as in Equation \eqref{eq15}.

\begin{equation}
\text{Rand}(C_1, C_2) = \frac{a + b}{n(n - 1)/2}
\label{eq15}
\end{equation}

The ARI is a corrected-for-chance version of the Rand Index. This correction uses the expected similarity of all comparisons between clusterings specified by a random model to set up a baseline. The ARI is computed as in Equation \eqref{eq16}.

\begin{equation}
\text{ARI}(C_1, C_2) = \frac{\text{Rand}(C_1, C_2) - \text{Expected Index}}{\text{Maximum Index} - \text{Expected Index}},
\label{eq16}
\end{equation}

\noindent where Maximum Index is expected to be 1 and Expected Index is the already mentioned expected degree of similarity with a random model. It is easy to see that $\text{ARI}(C_1, C_2) \in [-1,1]$, such that an ARI value close to 1 means a high degree of agreement between $C_1$ and $C_2$, a positive value close to 0 means no agreement and a value smaller that 0 means that the $\text{Rand}(C_1, C_2)$ is less than expected when comparing with random partitions. To summarize, the higher the ARI, the greater the degree of similarity between $C_1$ and $C_2$. For more details on ARI see \cite{hubert1985comparing}.

Our objective is to quantify the quality of the solutions obtained as a result of the methods presented in this paper. To accomplish this task we just set one of the two partitions given to compute ARI as the ground truth labels.

\subsection{Validation of results} \label{sec:ValidtnMethod}

In order to validate the results which will be presented in Section \ref{sec:results} we will use Bayesian statistical tests, instead of the classic Null Hypothesis Statistical Tests (NHST). In \cite{benavoli2017time} we find an in-depth analysis of the disadvantages of NHST, and a new model is proposed for carrying out comparisons researchers are interested in. \textit{"In a nutshell: NHST do not answer the question we ask"}. To put it clear, the disadvantages of the NHST that the authors highlight in \cite{benavoli2017time} are based on the trap of black-and-white thinking, this is: to reject, or not to reject?

To start with, NHST do not provide us with the probabilities associated to the analyzed hypotheses, and therefore it is not possible to answer the question: what is the probability that two methods are different? Another pitfall of NHST is that, with a sufficiently large number of observations, it is possible to reject almost any hypothesis. This is because the p-value does not allow us to separate between the effective size and the sample size, which is established by the researcher.

Also, NHST do not provide information about the magnitude of the effects and the uncertainty of its estimate. As a consequence, NHST may reject hypotheses despite very small effects, or even if there is significant uncertainty in the magnitude of the effects.

Furthermore, and this is a situation that all researchers have faced, NHST do not provide any information about the null hypothesis! That is: What can we conclude when NHST do not reject the null hypothesis? We can not infer anything since NHST can not provide evidence in its favor.

Finally, there are two other problems that researchers face when performing NHST. The first one is the choice of the significance level $\alpha$, for which there are no objective guidelines despite being critical to the test results. The second one is the need to previously formalize the intentions of the sampling of the results, which are usually fixed a posteriori; this could lead to a misreading of said results.

As shown in \cite{benavoli2017time}, most of these problems can be avoided by using Bayesian tests instead of NHST. In particular we will use the Bayesian sign test, which is the Bayesian version of the frequentist non-parametric sign test. To make use of it we will employ the R package \texttt{rNPBST}, whose documentation and guide can be found in \cite{carrasco2017rnpbst}.

We will use the Bayesian sign test to validate our results, which is based on obtaining the statistical distribution of a certain parameter $\rho$ according to the difference between the results, under the assumption that said distribution is a Dirichlet distribution. To get the distribution of $\rho$ we count the number of times that $A - B < 0$, the number of times where there are no significant differences, and the number of times that $A - B > 0$. In order to identify cases where there are no significant differences, we define the region of practical equivalence (rope) $[r_\text{min}, r_\text{max}]$, so that $P(A \approx B) = P(\rho \in \text{rope})$. Using these results we calculate the weights of the Dirichlet distribution and sample it to get a set of triplets with the following form: 
$$[P(\rho < r_\text{min}) = P(A - B < 0),\;\; P(\rho \in \text{rope}),\;\; P(\rho > r_\text{max}) = P(A - B > 0)]$$

\subsection{Calibration}

Table \ref{tab:params} shows a summary of the parameter setup used for the DILS$_{CC}$.

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{>{\centering\arraybackslash}c m{6cm} c}
			\hline
			Parameter & Meaning & Value \\
			\hline
			Evals & Fitness function evaluations & 300000 \\
			$p_s$ & Segment size for the mutation operator & $0.3 \times n$ \\
			$p_m$ & Maximum number of neighbors generated in each call to the LS procedure & Evals $\times 0.01$ \\
			$p_r$ & Probability that a feature is selected from $m_b$ for recombination & $0.3$ \\
			$\xi$ & Reinitialization method tolerance & $0.2$ \\
			$k$ & Output partition number of clusters & No. Classes (Table \ref{tab:datasets}) \\
			\hline

		\end{tabular}}
		%\end{adjustwidth}

	\caption{Parameters setup used for DILS$_{CC}$.}
	\label{tab:params}
\end{table}

The stop criterion is given by the number of evaluations of the fitness function $f$, which at most will be 300000. The implementation for DILS$_{CC}$ can be found at \href{https://github.com/GermangUgr/DILS_CC}{GitHub}.

To compare with the state-of-the-art methods mentioned in Section \ref{sec:BackSOTA} we will use the parameters setup shown in Table \ref{tab:paramsSOTA} for the implementation that can be found at \href{https://github.com/GermangUgr/TFG/tree/master/Software}{GitHub}. The implementation for the BRKGA algorithm can be found in the same link as DILS$_{CC}$. In all cases the value for $k$ is equal to the number of classes displayed in Table \ref{tab:datasets}.

\textcolor{red}{Meter todos los algoritmos en el mismo repo para no tener que poner tantos enlaces a github? Poner los enlaces como notas a pie de pagina tambien?}

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{cl}
			\hline
			Method name & Parameters name and values \\
			\hline
			BRKGA & Evals = $300000$; $|P| = 100$; $P_e = P_m = 0.2 * |P|$; $p_{\text{inherit}} = 50\%$ \\
			COPKM & \texttt{max\_iter} = $300$; \texttt{tolerance} = $1 * 10^{-4}$; \texttt{init\_mode} = \texttt{``rand''} \\
			CECM & \texttt{max\_iter} = $300$; $\alpha = 1$, $\rho = 100$, $\xi = 0.5$, \texttt{stop\_threshold} = $1 * 10^{-3}$, \texttt{init\_mode} = \texttt{``rand''} \\
			LCVQE & \texttt{max\_iter} = $300$; \texttt{initial\_centroids} = $\emptyset$ \\
			RDPM & \texttt{max\_iter} = $300$; $\xi_0 = 0.1$, $\xi_\text{rate} = 1$, $\lambda$ is calculated on the basis of the mean distances in the dataset. \\
			TVClust & \texttt{max\_iter} = $300$; $\alpha_0 = 1.2$, \texttt{stop\_threshold} = $5*10^{-4}$ \\
			\hline

		\end{tabular}}
		%\end{adjustwidth}

	\caption{Parameters setup used for the state-of-the-art algorithms.}
	\label{tab:paramsSOTA}
\end{table}

Parameter values have been assigned following the guidelines of the original creators of the different proposals. Since the evaluation in the experimental stage makes use of a high number of datasets, tuning each parameter specifically for each dataset is not feasible. Indeed, our goal is not optimization in a case-by-case basis but instead a comparison in the most general scenario possible. Therefore, given that the purpose of this work is to draw a fair comparison between the algorithms and assess their robustness in a common environment with multiple datasets, we have not included a tuning step to maximize any particular performance metric.

\section{Experimental Results} \label{sec:results}

In this section we present Tables from \ref{tab:results10} to \ref{tab:results20}, which display the results obtained by DILS$_{CC}$ and six state-of-the-art methods to be compared for each dataset and constraint set.

Since some of the methods we are comparing involve non-deterministic procedures, the results may vary from one run to another. To lessen the effect this may have on the results, we will apply each method 5 times to every dataset and constraint set, so that the measures shown in the previously mentioned tables correspond to the average of the 5 runs.

It should be noted that there are some missing results in these tables. In the case of the COPKM algorithm, this is due to the fact that it is highly dependent on the order in which constraints are analyzed. It is possible that COPKM cannot find a solution, even though it is always feasible, since the constraints have been generated based on the true labels. In the case of CECM, some of the results are not available because the memory structures that hold the algorithm grow non-linearly with the number of classes and the number of features of the dataset to be analyzed.

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & \textbf{0.611} & 0.022 & - & 0.335 & 0.316 & 0.025 & -0.005 \\
			Breast Cancer & \textbf{0.755} & 0.395 & -0.604 & 0.486 & 0.502 & 0.000 & 0.000 \\
			Bupa & \textbf{0.870} & 0.281 & - & -0.005 & -0.005 & -0.004 & -0.011 \\
			Circles & \textbf{0.798} & 0.166 & - & -0.003 & 0.162 & 0.137 & 0.133 \\
			Ecoli & 0.069 & 0.009 & - & 0.387 & \textbf{0.417} & 0.265 & - \\
			Glass & 0.009 & 0.016 & 0.184 & \textbf{0.268} & 0.197 & 0.211 & - \\
			Haberman & \textbf{0.638} & 0.047 & - & -0.002 & 0.127 & 0.218 & -0.004 \\
			Hayesroth & 0.031 & 0.027 & - & \textbf{0.106} & 0.097 & 0.054 & 0.139 \\
			Heart & \textbf{0.793} & 0.483 & - & 0.027 & 0.036 & 0.423 & -0.003 \\
			Ionosphere & \textbf{0.792} & 0.125 & - & 0.168 & 0.197 & 0.000 & 0.030 \\
			Iris & 0.621 & 0.238 & -0.285 & \textbf{0.730} & 0.607 & 0.244 & 0.684 \\
			Led7Digit & 0.016 & 0.009 & \textbf{0.497} & 0.425 & 0.369 & 0.316 & - \\
			Monk2 & 0.815 & 0.421 & \textbf{0.982} & 0.072 & 0.094 & -0.002 & 0.007 \\
			Moons & \textbf{0.968} & 0.277 & - & 0.241 & 0.319 & 0.785 & 0.092 \\
			Movement Libras & 0.022 & 0.004 & 0.285 & \textbf{0.293} & 0.256 & 0.000 & - \\
			Newthyroid & 0.028 & 0.017 & - & 0.579 & 0.289 & \textbf{0.846} & 0.002 \\
			Saheart & 0.800 & 0.147 & \textbf{0.974} & 0.018 & 0.020 & 0.068 & 0.000 \\
			Sonar & \textbf{0.649} & 0.124 & - & 0.004 & 0.013 & 0.000 & 0.000 \\
			Soybean & 0.304 & 0.496 & 0.503 & 0.545 & \textbf{0.621} & 0.000 & 0.244 \\
			Spectfheart & \textbf{0.896} & 0.334 & - & -0.107 & -0.114 & 0.000 & 0.050 \\
			Spiral & \textbf{0.856} & 0.058 & - & -0.003 & 0.012 & 0.034 & -0.002 \\
			Tae & \textbf{0.031} & 0.018 & - & 0.009 & -0.000 & 0.067 & 0.000 \\
			Vowel & 0.002 & 0.003 & - & 0.063 & -0.003 & \textbf{0.067} & - \\
			Wine & 0.334 & 0.120 & - & 0.360 & \textbf{0.368} & 0.281 & 0.004 \\
			Zoo & 0.151 & 0.105 & \textbf{0.715} & 0.666 & 0.412 & 0.335 & - \\
			\hline
			Mean & \textbf{0.474} & 0.158 & -0.510 & 0.226 & 0.212 & 0.175 & -0.186 \\
			\hline

		\end{tabular}}
		%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{10}$ by DILS$_{CC}$ and the state-of-the-art methods.}
	\label{tab:results10}
\end{table}

Table \ref{tab:results10} shows the results for the $SC_{10}$ constraint set. It is possible to notice that DILS$_{CC}$ represents a consistent improvement over all other 5 methods. This is due to the fact that the exploitation capacity of DILS$_{CC}$ allows it to take better advantage of the information contained in the constraint set than the rest of the methods, even with the lowest level of constraint-based information we work with.

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & \textbf{0.957} & 0.296 & - & 0.305 & 0.284 & 0.025 & -0.006 \\
			Breast Cancer & 0.788 & 0.772 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.037 \\
			Bupa & 0.984 & 0.840 & \textbf{1.000} & -0.005 & -0.007 & -0.007 & 0.001 \\
			Circles & \textbf{1.000} & 0.828 & \textbf{1.000} & -0.003 & 0.375 & 0.973 & 0.105 \\
			Ecoli & 0.134 & 0.015 & - & 0.387 & 0.372 & \textbf{0.719} & - \\
			Glass & 0.057 & 0.017 & - & \textbf{0.280} & 0.253 & 0.229 & - \\
			Haberman & \textbf{1.000} & 0.866 & \textbf{1.000} & -0.002 & 0.075 & 0.219 & -0.050 \\
			Hayesroth & \textbf{0.386} & 0.107 & - & 0.106 & 0.107 & 0.150 & 0.135 \\
			Heart & \textbf{1.000} & 0.692 & \textbf{1.000} & 0.025 & 0.033 & 0.442 & 0.000 \\
			Ionosphere & 0.970 & 0.803 & \textbf{1.000} & 0.178 & 0.212 & 0.000 & 0.057 \\
			Iris & \textbf{0.850} & 0.412 & - & 0.730 & 0.547 & 0.421 & 0.684 \\
			Led7Digit & 0.020 & 0.004 & - & 0.425 & \textbf{0.490} & 0.324 & - \\
			Monk2 & 0.902 & 0.751 & \textbf{1.000} & 0.072 & 0.170 & -0.002 & 0.007 \\
			Moons & \textbf{1.000} & 0.958 & \textbf{1.000} & 0.241 & 0.436 & 0.987 & 0.095 \\
			Movement Libras & 0.025 & 0.006 & -0.742 & \textbf{0.293} & 0.255 & 0.000 & - \\
			Newthyroid & 0.440 & 0.320 & - & 0.568 & 0.356 & \textbf{0.923} & 0.052 \\
			Saheart & 0.862 & 0.749 & \textbf{1.000} & 0.020 & 0.037 & 0.221 & 0.000 \\
			Sonar & \textbf{0.981} & 0.800 & - & 0.004 & 0.019 & 0.000 & 0.000 \\
			Soybean & 0.469 & 0.428 & 0.584 & 0.545 & \textbf{0.605} & 0.000 & 0.000 \\
			Spectfheart & \textbf{1.000} & 0.924 & 0.983 & -0.107 & -0.117 & 0.000 & -0.070 \\
			Spiral & \textbf{1.000} & 0.847 & - & -0.003 & 0.014 & 0.006 & 0.051 \\
			Tae & \textbf{0.299} & 0.119 & - & 0.008 & -0.004 & 0.062 & 0.000 \\
			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.073} & - \\
			Wine & \textbf{0.725} & 0.528 & - & 0.360 & 0.367 & 0.348 & -0.001 \\
			Zoo & 0.185 & 0.157 & 0.435 & \textbf{0.642} & 0.450 & 0.353 & - \\
			\hline
			Mean & \textbf{0.641} & 0.490 & -0.070 & 0.225 & 0.233 & 0.258 & -0.196 \\
			\hline

		\end{tabular}}
		%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{15}$ by DILS$_{CC}$ and the state-of-the-art methods.}
	\label{tab:results15}
\end{table}

The results obtained with the $SC_{15}$ constraint set are presented in Table \ref{tab:results15}. We observe how DILS$_{CC}$ and COPKM are able to obtain ARI values equal to 1, meaning that the resulting partitions of the datasets perfectly match the true partitions. The fact that DILS$_{CC}$ is always able to output a partition of the datasets is a noteworthy advantage over COPKM. Even in the cases where COPKM is unable to produce a partition, DILS often finds the optimal or a near-optimal partition.

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.4}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{lccccccc}
			\hline
			Dataset & DILS$_{CC}$ & BRKGA & COPKM & LCVQE & RDPM & TVClust & CEKM \\
			\hline
			Appendicitis & \textbf{1.000} & \textbf{1.000} & - & 0.305 & 0.331 & 0.012 & -0.006 \\
			Breast Cancer & 0.807 & 0.807 & \textbf{1.000} & 0.486 & 0.502 & 0.000 & 0.000 \\
			Bupa & 0.991 & 0.936 & \textbf{1.000} & -0.005 & -0.007 & -0.006 & 0.000 \\
			Circles & \textbf{1.000} & 0.935 & \textbf{1.000} & -0.003 & 0.629 & 1.000 & 0.138 \\
			Ecoli & 0.201 & 0.028 & - & 0.387 & 0.459 & \textbf{0.763} & - \\
			Glass & \textbf{0.322} & 0.038 & - & 0.271 & 0.287 & 0.218 & - \\
			Haberman & \textbf{1.000} & 0.943 & \textbf{1.000} & -0.002 & 0.106 & 0.737 & 0.000 \\
			Hayesroth & \textbf{0.812} & 0.548 & - & 0.106 & 0.107 & 0.208 & 0.056 \\
			Heart & \textbf{1.000} & 0.942 & \textbf{1.000} & 0.025 & 0.033 & 1.000 & 0.000 \\
			Ionosphere & 0.982 & 0.902 & \textbf{1.000} & 0.173 & 0.309 & 0.000 & 0.115 \\
			Iris & \textbf{0.968} & 0.691 & - & 0.708 & 0.540 & 0.585 & 0.684 \\
			Led7Digit & 0.017 & 0.007 & - & 0.425 & \textbf{0.571} & 0.327 & - \\
			Monk2 & 0.904 & 0.855 & \textbf{1.000} & 0.072 & 0.253 & -0.002 & 0.160 \\
			Moons & \textbf{1.000} & 0.937 & \textbf{1.000} & 0.241 & 0.831 & 1.000 & 0.180 \\
			Movement Libras & 0.018 & 0.006 & - & \textbf{0.293} & \textbf{0.293} & 0.000 & - \\
			Newthyroid & 0.842 & 0.848 & - & 0.568 & 0.393 & \textbf{0.933} & 0.004 \\
			Saheart & 0.881 & 0.881 & \textbf{1.000} & 0.018 & 0.026 & \textbf{1.000} & -0.006 \\
			Sonar & \textbf{1.000} & 0.992 & \textbf{1.000} & 0.004 & 0.127 & 0.000 & 0.003 \\
			Soybean & 0.564 & 0.416 & -0.218 & 0.545 & \textbf{0.631} & 0.000 & -0.016 \\
			Spectfheart & \textbf{1.000} & 0.953 & \textbf{1.000} & -0.107 & -0.112 & 0.000 & -0.054 \\
			Spiral & \textbf{1.000} & 0.921 & \textbf{1.000} & -0.003 & 0.011 & 0.006 & 0.045 \\
			Tae & \textbf{0.892} & 0.351 & - & 0.008 & 0.000 & 0.035 & 0.000 \\
			Vowel & 0.003 & 0.002 & - & 0.063 & -0.003 & \textbf{0.071} & - \\
			Wine & 0.916 & \textbf{0.937} & - & 0.360 & 0.363 & 0.380 & 0.015 \\
			Zoo & 0.203 & 0.176 & \textbf{0.821} & 0.642 & 0.439 & 0.335 & - \\
			\hline
			Mean & \textbf{0.733} & 0.642 & 0.064 & 0.223 & 0.285 & 0.344 & -0.187 \\
			\hline

		\end{tabular}}
		%\end{adjustwidth}

	\caption{Experimental results obtained for $CS_{20}$ by DILS$_{CC}$ and the state-of-the-art methods.}
	\label{tab:results20}
\end{table}

In Table \ref{tab:results15}, which shows the results obtained with the $SC_{15}$ constraint set, we can see how the quality of the results of DILS$_{CC}$ scales with the number of constraints, as well as that of methods such as BRKGA or COPKM (the latter being more limited in other aspects). It is also worth noting how DILS$_{CC}$ provides the best average ARI for each of the three constraint sets analyzed in this paper.

\clearpage

\section{Statistical Analysis of Results} \label{sec:analisis}

With the results obtained by all methods for a total of 75 different datasets---the 25 datasets in combination with the 3 constraint sets for each one of them---we can perform an empirical analysis of them. This way we can statistically determine whether DILS$_{CC}$ represents a significant improvement over previous proposals.

Having in mind the notation introduced in Section \ref{sec:ValidtnMethod}, and as far as the comparison of population-based methods is concerned, we will refer to the results obtained by a given state-of-the-art method as sample $A$, and to the results obtained with DILS$_{CC}$ as sample $B$.

One of the major advantages of the Bayesian sign test is that we can obtain a very illustrative visual representation of its results. We can produce a representation of the triplet set in the form of a heatmap where each triplet constitutes one point whose location is given by barycentric coordinates. With this in mind, we will associate each of the triplet values with each of the three vertices of an equilateral triangle. In order to find out where a certain triplet will be placed within the triangle, we will take each of its three values and draw a parallel line to the opposing side of the corresponding vertex; the separation between a triangle side and its parallel line will be proportional to the associated triplet value, so that the higher the value, the closer the line will be to the vertex. The location where the three lines intersect is where we draw a point. Since the values of every triplet describe a probability distribution and therefore they must add up to one, we can be sure that all triplets will lie in some point within the triangle. The color indicates the density of points in a given region, with yellow representing a high density and red a low density.

\begin{figure}[ht!]
	\centering
	\begin{subfigure}{.45\textwidth}
		\includegraphics[width=1\linewidth]{Figures/DILSvsBRKGA.pdf}
		\caption{BRKGA}
		\label{fig:DILSvsBRKGA}
	\end{subfigure}
	\begin{subfigure}{.45\textwidth}
		\includegraphics[width=1\linewidth]{Figures/DILSvsCOPKM.pdf}
		\caption{COPKM}
		\label{fig:DILSvsCOPKM}
	\end{subfigure}
	\begin{subfigure}{.45\textwidth}
		\includegraphics[width=1\linewidth]{Figures/DILSvsLCVQE.pdf}
		\caption{LCVQE}
		\label{fig:DILSvsLCVQE}
	\end{subfigure}
	\begin{subfigure}{.45\textwidth}
		\includegraphics[width=1\linewidth]{Figures/DILSvsTVClust.pdf}
		\caption{TVClust}
		\label{fig:DILSvsTVClust}
	\end{subfigure}
	\begin{subfigure}{.45\textwidth}
		\includegraphics[width=1\linewidth]{Figures/DILSvsRDPM.pdf}
		\caption{RDPM}
		\label{fig:DILSvsRDPM}
	\end{subfigure}
	\begin{subfigure}{.45\textwidth}
		\includegraphics[width=1\linewidth]{Figures/DILSvsCEKM.pdf}
		\caption{CEKM}
		\label{fig:DILSvsCEKM}
	\end{subfigure}
	\caption{Heat diagrams comparing DILS with the state-of-the-art methods.}
	\label{fig:heatDiag}
\end{figure}

Figure \ref{fig:heatDiag} shows all six heatmaps obtained when applying the Bayesian sign test to DILS$_{CC}$ and the state-of-the-art methods. All heatmaps suggest that the test assigns a high probability to $A - B < 0$, since most triplets are presented in the lower left third of the triangle. Bearing this in mind, and considering that ARI is a measure to maximize, we can say that the Bayesian sign test provides strong evidence that DILS$_{CC}$ does represent a significant improvement over all other six methods, being the results offered by the latter non equivalent to those offered by DILS$_{CC}$ and lower in quality.

\clearpage

\section{Conclusions} \label{sec:conclusiones}

In this paper we proposed the DILS heuristic method, along with its application to the SSL constrained clustering problem, which we called DILS$_{CC}$. Focusing on instance-level ML and CL constraints, DILS$_{CC}$ has proven that heuristic techniques constitute a competitive approach to constrained clustering, being able to scale the quality of the results in a directly proportional way to the number of constraints.

Supported by the Bayesian statistical tests, we were able to objectively prove that the DILS$_{CC}$ approach is significantly better than the state-of-the-art methods. DILS$_{CC}$ has proven to be better when considering the quality and availability of the solutions, especially in cases where large sets of constraints are analyzed.

\section*{Acknowledgements}

Our work has been supported by the research project TIN2017-89517-P and PP2016.PRI.I.02.

\section*{Compliance with ethical standards}

\section*{Declaration of competing interest}

The authors declare that there is no conflict of interest.

\clearpage

\section*{References}

\bibliography{mybibfile}

\end{document}