### 2019-07-06 Memoria/FrontBackmatter/Abstract.tex

\pdfbookmark[1]{Abstract}{Abstract}
\begingroup

\chapter*{Abstract}

Clustering has always been a powerful tool in knowledge discovery. Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning: constrained clustering. This technique is a generalization of traditional clustering that considers additional information encoded by constraints. Constraints can be given in the form of instance-level must-link and cannot-link constraints, which this paper focuses on. We propose the first application of Differential Evolution to the constrained clustering problem, as well as a new Local Search variant and its application to the problem at hand. We will compare the results obtained by these new proposals to those obtained by the state-of-the-art algorithms on twenty datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.\\

\noindent\textbf{Keywords:} constrained clustering, instance-level, must-link, cannot-link, Differential Evolution, Local Search.

\endgroup